---
title: "Featuring Engineering Practice"
output: 
  html_document:
    toc: true
    toc_depth: 3
author: Machine Learning II
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# setwd("C:\\Users\\angel\\Dropbox\\IE\\1st Assignement\\house-prices")
Sys.setenv(JAVA_HOME='C:\\Program Files\\Java\\jre-9.0.4')
library(ggplot2)
library(plyr)
library(dplyr)
library(moments)
library(glmnet)
library(caret)
library(FSelector)
```

# Introduction
In this first practical session we will make the first contact with the featuring engineering process and its impact in a ML pipeline.
Feature engineering is a very important part of the process of developing prediction models. It is considered, by many authors, an art, and it involves human-driven design and intuition. Feature engineering sessions will try to uncover the most relevant issues that must be addressed, and also provide some guidelines to start building sound feature engineering processes for ML problems. 

The experimental dataset we are going to use is the House Prices Dataset. It includes 79 explanatory variables of residential homes. For more details on the dataset and the competition see <https://www.kaggle.com/c/house-prices-advanced-regression-techniques>.

## What is my goal?
- I have to predict predict the final price of each home (Therefore, this is a regression task)
- I have to use the feature engineering techniques explained in class to transform the dataset.

# Data Reading and preparation
The dataset is offered in two separated fields, one for the training and another one for the test set. Best way of moving forward is to sign in into Kaggle and download them from the 'Data' section.

```{r Load Data}
training_data = read.csv(file = file.path("data", "train.csv"))
test_data = read.csv(file = file.path("data", "test.csv"))
```

```{r Check for duplicates}
length(unique(training_data$Id)) == nrow(training_data)
```

There is no duplicates so we remove the Id column
```{r Remove the ID Column}
training_data = training_data[ , -which(names(training_data) %in% c("Id"))]

```

## Hunting NAs
Our dataset is filled with many missing values, therefore, before we can build any predictive model we'll clean our data by filling in all NA's with appropriate values.

Counting columns with null values
```{r NAs discovery}
na.cols <- which(colSums(is.na(training_data)) > 0)
sort(colSums(sapply(training_data[na.cols], is.na)), decreasing = TRUE)
paste('There are', length(na.cols), 'columns with missing values')
```

NA imputation:
```{r Train NA Imputation}
# Alley : NA means "no alley access"
training_data$Alley = factor(training_data$Alley, levels=c(levels(training_data$Alley), "None"))
training_data$Alley[is.na(training_data$Alley)] = "None"

training_data$BedroomAbvGr[is.na(training_data$BedroomAbvGr)] <- 0

# Bsmt : NA for basement features is "no basement"
training_data$BsmtQual = factor(training_data$BsmtQual, levels=c(levels(training_data$BsmtQual), "No"))
training_data$BsmtQual[is.na(training_data$BsmtQual)] = "No"

training_data$BsmtCond = factor(training_data$BsmtCond, levels=c(levels(training_data$BsmtCond), "No"))
training_data$BsmtCond[is.na(training_data$BsmtCond)] = "No"

training_data$BsmtExposure[is.na(training_data$BsmtExposure)] = "No"

training_data$BsmtFinType1 = factor(training_data$BsmtFinType1, levels=c(levels(training_data$BsmtFinType1), "No"))
training_data$BsmtFinType1[is.na(training_data$BsmtFinType1)] = "No"

training_data$BsmtFinType2 = factor(training_data$BsmtFinType2, levels=c(levels(training_data$BsmtFinType2), "No"))
training_data$BsmtFinType2[is.na(training_data$BsmtFinType2)] = "No"

# Fence : NA means "no fence"
training_data$Fence = factor(training_data$Fence, levels=c(levels(training_data$Fence), "No"))
training_data$Fence[is.na(training_data$Fence)] = "No"

# FireplaceQu : NA means "no fireplace"
training_data$FireplaceQu = factor(training_data$FireplaceQu, levels=c(levels(training_data$FireplaceQu), "No"))
training_data$FireplaceQu[is.na(training_data$FireplaceQu)] = "No"

# Garage : NA for garage features is "no garage"
training_data$GarageType = factor(training_data$GarageType, levels=c(levels(training_data$GarageType), "No"))
training_data$GarageType[is.na(training_data$GarageType)] = "No"

training_data$GarageFinish = factor(training_data$GarageFinish, levels=c(levels(training_data$GarageFinish), "No"))
training_data$GarageFinish[is.na(training_data$GarageFinish)] = "No"

training_data$GarageQual = factor(training_data$GarageQual, levels=c(levels(training_data$GarageQual), "No"))
training_data$GarageQual[is.na(training_data$GarageQual)] = "No"

training_data$GarageCond = factor(training_data$GarageCond, levels=c(levels(training_data$GarageCond), "No"))
training_data$GarageCond[is.na(training_data$GarageCond)] = "No"

# LotFrontage : NA most likely means no lot frontage
training_data$LotFrontage[is.na(training_data$LotFrontage)] <- 0

# MasVnrType : NA most likely means no veneer
training_data$MasVnrType[is.na(training_data$MasVnrType)] = "None"
training_data$MasVnrArea[is.na(training_data$MasVnrArea)] <- 0

# MiscFeature : NA = "no misc feature"
training_data$MiscFeature = factor(training_data$MiscFeature, levels=c(levels(training_data$MiscFeature), "No"))
training_data$MiscFeature[is.na(training_data$MiscFeature)] = "No"

# PoolQC : data description says NA means "no pool"
training_data$PoolQC = factor(training_data$PoolQC, levels=c(levels(training_data$PoolQC), "No"))
training_data$PoolQC[is.na(training_data$PoolQC)] = "No"

# Electrical : NA means "UNK"
training_data$Electrical = factor(training_data$Electrical, levels=c(levels(training_data$Electrical), "UNK"))
training_data$Electrical[is.na(training_data$Electrical)] = "UNK"

# GarageYrBlt: It seems reasonable that most houses would build a garage when the house itself was built.
idx <- which(is.na(training_data$GarageYrBlt))
training_data[idx, 'GarageYrBlt'] <- training_data[idx, 'YearBuilt']

na.cols <- which(colSums(is.na(training_data)) > 0)
paste('There are now', length(na.cols), 'columns with missing values')
```

We repeat the process for test_data
```{r Test Inputation}
# Alley : data description says NA means "no alley access"
test_data$Alley = factor(test_data$Alley, levels=c(levels(test_data$Alley), "None"))
test_data$Alley[is.na(test_data$Alley)] = "None"

test_data$BedroomAbvGr[is.na(test_data$BedroomAbvGr)] <- 0

# BsmtQual etc : data description says NA for basement features is "no basement"
test_data$BsmtQual = factor(test_data$BsmtQual, levels=c(levels(test_data$BsmtQual), "No"))
test_data$BsmtQual[is.na(test_data$BsmtQual)] = "No"

test_data$BsmtCond = factor(test_data$BsmtCond, levels=c(levels(test_data$BsmtCond), "No"))
test_data$BsmtCond[is.na(test_data$BsmtCond)] = "No"

test_data$BsmtExposure[is.na(test_data$BsmtExposure)] = "No"

test_data$BsmtFinType1 = factor(test_data$BsmtFinType1, levels=c(levels(test_data$BsmtFinType1), "No"))
test_data$BsmtFinType1[is.na(test_data$BsmtFinType1)] = "No"

test_data$BsmtFinType2 = factor(test_data$BsmtFinType2, levels=c(levels(test_data$BsmtFinType2), "No"))
test_data$BsmtFinType2[is.na(test_data$BsmtFinType2)] = "No"

# Fence : data description says NA means "no fence"
test_data$Fence = factor(test_data$Fence, levels=c(levels(test_data$Fence), "No"))
test_data$Fence[is.na(test_data$Fence)] = "No"

# FireplaceQu : data description says NA means "no fireplace"
test_data$FireplaceQu = factor(test_data$FireplaceQu, levels=c(levels(test_data$FireplaceQu), "No"))
test_data$FireplaceQu[is.na(test_data$FireplaceQu)] = "No"

# GarageType etc : data description says NA for garage features is "no garage"
test_data$GarageType = factor(test_data$GarageType, levels=c(levels(test_data$GarageType), "No"))
test_data$GarageType[is.na(test_data$GarageType)] = "No"

test_data$GarageFinish = factor(test_data$GarageFinish, levels=c(levels(test_data$GarageFinish), "No"))
test_data$GarageFinish[is.na(test_data$GarageFinish)] = "No"

test_data$GarageQual = factor(test_data$GarageQual, levels=c(levels(test_data$GarageQual), "No"))
test_data$GarageQual[is.na(test_data$GarageQual)] = "No"

test_data$GarageCond = factor(test_data$GarageCond, levels=c(levels(test_data$GarageCond), "No"))
test_data$GarageCond[is.na(test_data$GarageCond)] = "No"

# LotFrontage : NA most likely means no lot frontage
test_data$LotFrontage[is.na(test_data$LotFrontage)] <- 0

# MasVnrType : NA most likely means no veneer
test_data$MasVnrType[is.na(test_data$MasVnrType)] = "None"
test_data$MasVnrArea[is.na(test_data$MasVnrArea)] <- 0

# MiscFeature : data description says NA means "no misc feature"
test_data$MiscFeature = factor(test_data$MiscFeature, levels=c(levels(test_data$MiscFeature), "No"))
test_data$MiscFeature[is.na(test_data$MiscFeature)] = "No"

# PoolQC : data description says NA means "no pool"
test_data$PoolQC = factor(test_data$PoolQC, levels=c(levels(test_data$PoolQC), "No"))
test_data$PoolQC[is.na(test_data$PoolQC)] = "No"

```

## Factorize features
Some numerical features are actually really categories. Therefore we transform the feature from numeric to categorical
```{r Factorize features}

training_data$MSSubClass <- as.factor(training_data$MSSubClass)
training_data$MoSold <- as.factor(training_data$MoSold)

test_data$MSSubClass <- as.factor(test_data$MSSubClass)
test_data$MoSold <- as.factor(test_data$MoSold)
```


## Skewness
If we print the histogram of the target value, we obseve a large skewness in the Target value (i.e., the distribution in not normally distributed).
To solve that we log transform this variable so that it becomes normally distributed. A normally distributed target variable helps in the modeling step (i.e., the finding of the relationship between target and independent variables).
```{r}
# get data frame of SalePrice and log(SalePrice + 1) for plotting
df <- rbind(data.frame(version="log(price+1)",x=log(training_data$SalePrice + 1)),
            data.frame(version="price",x=training_data$SalePrice))

ggplot(data=df) +
  facet_wrap(~version,ncol=2,scales="free_x") +
  geom_histogram(aes(x=x), bins = 50)
```


We therefore transform the target value applying log.
```{r Log transform the target for official scoring}
# Log transform the target for official scoring
training_data$SalePrice <- log1p(training_data$SalePrice)
```


The same "skewness" observed in the target variable also affects other variables. To facilitate the application of the regression model we are going to also eliminate this skewness
For numeric feature with excessive skewness, perform log transformation.
```{r}

column_types <- sapply(names(training_data),function(x){class(training_data[[x]])})
numeric_columns <-names(column_types[column_types != "factor"])

# skew of each variable
skew <- sapply(numeric_columns,function(x){skewness(training_data[[x]],na.rm = T)})

# transform all variables above a threshold skewness.
skew <- skew[skew > 0.75]
for(x in names(skew)) {
  training_data[[x]] <- log(training_data[[x]] + 1)
}
```

The same for the test data
```{r}
column_types <- sapply(names(test_data),function(x){class(test_data[[x]])})
numeric_columns <-names(column_types[column_types != "factor"])

skew <- sapply(numeric_columns,function(x){skewness(test_data[[x]],na.rm = T)})
skew <- skew[skew > 0.75]
for(x in names(skew)) {
  test_data[[x]] <- log(test_data[[x]] + 1)
}
```

## Train, Validation Spliting
We are going to split the annotated dataset in training and validation for the later evaluation of our regression models
```{r Train test split}
# I found this function, that is worth to save for future ocasions.
splitdf <- function(dataframe, seed=NULL) {
  if (!is.null(seed)) set.seed(seed)
 	index <- 1:nrow(dataframe)
 	trainindex <- sample(index, trunc(length(index)/1.5))
 	trainset <- dataframe[trainindex, ]
 	testset <- dataframe[-trainindex, ]
 	list(trainset=trainset,testset=testset)
}
splits <- splitdf(training_data, seed=1)
training <- splits$trainset
validation <- splits$testset
```

If we inspect in detail the categorical variables of the dataset, we'll see that some are incomplete: they only have a unique value for all the dataset.
These features are not valuable. Remember the three aspects that a feature should have: informative, <b>discriminative</b> and independent. Incomplete cases are not discriminative at all.
In addition, this might create problems when fitting the regression model

The following code show the incomplete cases
```{r}
## remove incomplete cases
paste("Training set incomplete cases")
sapply(lapply(na.omit(training)[sapply(na.omit(training), is.factor)], droplevels), nlevels)
paste("Validation set incomplete cases")
sapply(lapply(na.omit(validation)[sapply(na.omit(validation), is.factor)], droplevels), nlevels)
paste("Test set incomplete cases")
sapply(lapply(na.omit(test_data)[sapply(na.omit(test_data), is.factor)], droplevels), nlevels)
```

We remove the detected incomplete cases
```{r}
# Remove the Utilities feature from the dataset (It only has one value)
training <- training[,-which(names(training) == "Utilities")]
validation <- validation[,-which(names(validation) == "Utilities")]
test_data <- test_data[,-which(names(test_data) == "Utilities")]

```

# Feature Engineering
We here start the Feature Engineering.

## Filtering Methods
We will rank the features according to their predictive power according to the methodologies seen in class: the Chi Squared Independence test and the Information Gain.

#### Full Model
We must always fit a lm model with all the features to have a baseline to evaluate the impact of the feature engineering.
```{r Full Regression model, message=FALSE, warning=FALSE}

set.seed(121)
train_control_config <- trainControl(method = "repeatedcv", 
                       number = 5, 
                       repeats = 1,
                       returnResamp = "all")

full.lm.mod <- train(SalePrice ~ ., data = training, 
               method = "lm", 
               metric = "RMSE",
               preProc = c("center", "scale"),
               trControl=train_control_config)

for (x in names(validation)) {
  full.lm.mod$xlevels[[x]] <- union(full.lm.mod$xlevels[[x]], levels(validation[[x]]))
}
full.lm.mod.pred <- predict(full.lm.mod, validation[,-ncol(validation)])
full.lm.mod.pred[is.na(full.lm.mod.pred)] <- 0

my_data=as.data.frame(cbind(predicted=full.lm.mod.pred,observed=validation$SalePrice))

ggplot(my_data,aes(predicted,observed))+
  geom_point() + geom_smooth(method = "lm") +
  labs(x="Predicted") +
  ggtitle('Linear Model')

paste("Full Linear Regression RMSE = ", sqrt(mean((full.lm.mod.pred - validation$SalePrice)^2)))
```

### Chi-squared Selection
Making use of the `FSelector` package <https://cran.r-project.org/web/packages/FSelector/FSelector.pdf>, rank the features according to the Chi Squared value. If you've problems with this package (some of us have problems with it), do some research to find another packages that will provide the Chi squared selection.

Does it make sense to remove some features? Is so, do it! <b>(Tip: Sure it does)</b>
```{r}
# your code here
```

#### Evaluation
Evaluate the impact (in terms of RMSE) of the feature selection.
To that end, execute the previous LM model taking as input the filtered training set
```{r Chi-Squared Regression, message=FALSE, warning=FALSE}
# Compute your model with the new features and compute the new RMSE to decide whether to keep the new variables or not.
```

### Information Gain Selection
Let's experiment now with Information Gain Selection.
Making also use of the `FSelector` package <https://cran.r-project.org/web/packages/FSelector/FSelector.pdf>, rank the features according to their Information Gain and filter those which you consider, according to the IG value. 

Again, there're more alternatives to compute the IG.
```{r}
# your code here.
```

#### Evaluation
Evaluate the impact of the IG selection in the model performance
```{r Information Gain Regression Model, message=FALSE, warning=FALSE}
# Compute the new RMSE and decide.
```


### Exercise
Experiment with diferent cutoffs to filter the features and evaluate their impact in the perfomarnce. Select the cutoff that you consider more appropiate.
Using the result of the evaluation, filter the dataset (according to the method and cutoff that you decide)
```{r}

# Your code here

```

## Wrapper Methods
Let us experiment now with Wrapper Methods. In particular, we are going to apply Forward and Backward Stepwise Selection Methods to find the best feature combination for this dataset.

### Stepwise


#### Backward Stepwise
`caret` package provides a useful and easy way of experimenting with stepwise selection. Try it to know what a wrapper method suggests as the best possible subset of features and compare your results with the baseline.

```{r Backward Stepwise, message=FALSE, warning=FALSE}

# Your code here

```

Printout only the selected features.
```{r}
# Your code here
```

Comput the RMSE of the selected model
```{r}

# Compute RMSE

```

#### Forward Stepwise

Try the same with forward stepwise.

```{r Forward Stepwise, message=FALSE, warning=FALSE}

# Your code here

```

Printout only the selected features.
```{r}
# Your code here
```

Compute the new RMSE
```{r}

# Your code here

```


## Embedded
Finally, we will experiment with embedded methods. 
In particular we are going to focus on Ridge and  Lasso Regularization.

### Ridge Regression
For this exercise, we are going to make use of the <a href="https://cran.r-project.org/web/packages/glmnet/index.html">`glmnet`</a> library. Take a look to the library and fit a glmnet model for Ridge Regression, using the grid of lambda values provided.
```{r Ridge Regression, warning=FALSE}
lambdas <- 10^seq(-2, 3, by = .1)
ridge.mod <- glmnet(x = data.matrix(training[,-ncol(training)]), y=training$SalePrice, alpha = 0, lambda = lambdas)
```

#### Evaluation
Plotting the RMSE for the different lambda values, we can see the impact of this parameter in the model performance.
Small values seem to work better for this dataset
```{r Ridge Evaluation}
RMSE = numeric(length(lambdas))
for (i in seq_along(lambdas)){
  ridge.pred=predict(ridge.mod, s=lambdas[i], data.matrix(validation[,-ncol(validation)]))
  RMSE[i] <- sqrt(mean((ridge.pred - validation$SalePrice)^2))
}
plot(lambdas, RMSE, main="Ridge", log="x", type = "b")
```


##### Cross Validation
Making use of cv.glmnet <https://www.rdocumentation.org/packages/glmnet/versions/2.0-12/topics/cv.glmnet>, create a cross-validated Ridge Regression Model for the provided lambdas.

Plotting again the error, CV give us a better understanding on the impact of lambda in the model performance
```{r}
ridge.cv_fit <- cv.glmnet(x = data.matrix(training[,-ncol(training)]), y=training$SalePrice, alpha = 0, lambda = lambdas)
plot(ridge.cv_fit)
```


<b>Interpretation:</b>

1. The plot shows the MSE (red dots) for the provided lambda values (included in the grid).
2. The confidence intervals represent error estimates for the RSE, computed using CV. 
3. The vertical lines show the locations of lambda.min (lambda that achives the best MSE) and lambda.1se (the largest lambda value within 1 standard error of lambda.min. Using lambda.1se hedges against overfitting by selecting a larger lambda value than the min).
4. The numbers across the top are the number of nonzero coefficient estimates.


Select the best lambda form the CV model, use it to predict the target value of the validation set and evaluate the results (in terms of RMSE)
```{r}
# Your code here
```


Predict with that lambda on the validation set
```{r}
# Your code here
```

Let's plot the predictions against the actual values to have an idea of the model performance
```{r}
# Plot important coefficients
```

Rank the variables according to the importance attributed by the model
```{r}
# Your code here
```

### Lasso Regresion
Using again the <a href="https://cran.r-project.org/web/packages/glmnet/index.html">`glmnet`</a> library, fit a Lasso Regression (take a look to the alpha parameter) using the grid of lambda values provided.

Repeat now what you did with Ridge Regression with the Lasso method.
```{r}
# Your code here
```

